# Missing values

```{r}
#install.packages("covidcast")
library(covidcast)
library(tidyverse)
library(patchwork)
```

## Doctor Visit Data
```{r}
source('plot_missing.R')
```
```{r}
doctor_visit_data_raw <- covidcast_signal("doctor-visits", "smoothed_cli", start_day = "2020-02-01", end_day = "2021-02-01", geo_type='state',geo_values=c('ny','ca','tx','co','hi','wa','ks','va','mo'))
```

```{r}
plot_missing(doctor_visit_data_raw)
```

In the raw data, we found that there are two columns (sample_size & stderr) that consist of only missing values, and the other columns don't have missing values. 
The following is the number of missing values in each column:
```{r,warning=F}
colSums(is.na(doctor_visit_data_raw)) %>%
  sort(decreasing = TRUE)
```
Since the two columns have only missing values, we could just discard these two columns just as we discard the columns that consist of the same value all the way through.

## Hospital Admission Data
```{r}
hospital_data_raw <- covidcast_signal("hospital-admissions", "smoothed_covid19_from_claims", start_day = "2020-02-01", end_day = "2021-02-01", geo_type='state',geo_values=c('ny','ca','tx','co','hi','wa','ks','va','mo'))
```
```{r}
plot_missing(hospital_data_raw)
```

```{r,warning=F}
colSums(is.na(hospital_data_raw)) %>%
  sort(decreasing = TRUE)
```

Same situation here, so we just removed the two columns consisting of only missing values.

## The other datasets
Actually the same situation for the other datasets, so we just removed the columns consisting of missing values.
